NodIO: a framework and architecture for pool-based evolutionary computation

### 2016-05-03 EvoSoft16/nodio-evosoft16.tex
Solving hyphenation problems. References #125

@@

Let us see how this system addresses the different challenges
outlined above:
\begin{itemize}
\item {\em Scalability} is provided via the use of a lightweight and
  high-performance, single-threaded, server based in Node.js and
  Express.js. Although this single server is a bottleneck since it
  will eventually saturate, the fact that it runs as a non-blocking single thread
  allows the service of many requests. In fact, a limit in the
  number of simultaneous requests will be reached, but so far it has
  not been found, unlike what we found in our previous systems, DCoR,
  \cite{gecco07:workshop:dcor}, which had a low scalability. 
\item The framework is {\em heterogeneous} since it does not need any
  performance, operating system, or even browser requirement as long
  as JavaScript is enabled: anyone
  visiting the page, even from mobile devices, can load the algorithm.
  In case the browser does not support HTML5 Web workers (e.g., 
  Windows Phone or some version of Android) or their use
  has been disabled, a basic version of {\sf NodIO} can also be used.
\item {\em Fault tolerance} is always an issue, and in this case, the
  single point of failure would be the server: the system, as such,
  would break down if the server fails. However, the individual
  islands in every browser would continue running, and having access
  to just one of them would allow the local algorithm to proceed. In
  fact, the island does not need the server to run: it runs locally if
  needed, with the only exception that it is obviously unable to
  communicate with the rest of the islands.
\item {\em Adaptiveness} is achieved simply through the autonomous
  operation of every individual island without any synchronization
  mechanism. The islands in the system are, in fact, unaware of each
  other, communicating only through the server. The webpage itself can
  be displayed, using responsive design, in either desktop or mobile
  systems. 
\item Since the algorithm runs in the browser, user's {\em safety} is
  achieved through its sandbox mechanisms. The user is thus assured
  that there is no unsafe access either to their local files or even
  to more resources than the browser should be allotted.
\item Running the algorithm is just a matter of loading the page,
  which makes the operation totally {\em anonymous}. For the same
  reason, {\em ease of use} is optimal, being as easy as simply
  clicking on an URL, available to anyone with access to a browser.
\item {\em Reasonable performance} is not ensured. In fact, we should
  make sure that there is a reasonable amount of clients over which
  the performance achieved is better than what you would obtain in
  your own desktop system. If this is not the case, it is a pointless
  academic exercise.
\end{itemize}

@@

\begin{itemize}
\item {\em Randomize} Both Matlab and Java programs rely on a 
Java Randomizer library for the 
randomization of the shift vectors and matrices used by the function.
The library includes the generation of pseudorandom Gaussians of type
double and uniformly distributed integers. Generators use a single
{\tt long} seed. In order to implement the same functionality in
JavaScript, the library {\tt random-js} was used, as there are inconsistencies 
in the implementation of the standard {\tt Math.random()} between engines
and also their results are non-deterministic. 
The library {\tt random-js}, implements a Mersenne Twister algorithm producing 
consistent results across all JavaScript VMs.
\item {\em Timing functions} In order to compare the performance of 
each implementation, an accurate and consistent way of measuring 
the runtime of functions is needed.
In JavaScript the {\tt Date class} is often used 
to measure execution time, but has the limitation of having a maximum 
resolution given in milliseconds.
In order to measure the runtime intervals with a higher precision two functions
where used: for Node.js we used the native {\tt process.hrtime()} function
which returns a high-resolution measure in a {\tt [seconds, nanoseconds]}
array, and for measuring in the browser the {\tt Performance.now()} function was called,
which uses floating-point numbers with up to microsecond precision. 
A drawback found is that the {\tt Performance.now()} function is only implemented
in Firefox and Chrome browsers. Both functions are independent of the system clock. 
\item {\em Tools} An important consideration when choosing certain 
language to write non-trivial programs is the availability of development tools. While developing
the experiments, we used a comprehensive set of developer tools available for both
browsers and desktop \cite{tilkov2010node}. These tools included
package managers, in this case {\tt npm}, debuggers, logging libraries
such as {\tt winston}, 
network monitors, editors, and even debugging of the multi-threaded execution of Web Workers.
Although the fact that developer tools vary from browser to browser
supposed a drawback, they all have the same functionality. The
JavaScript debugger can be accessed directly from the user menu in
most browsers. 
\item {\em Data Types} JavaScript uses floating point numbers with a 
limited precision of 64 bits that partially implement the functionality of
the {\tt StrictMath} library used by the Java implementation. If more precision
is needed, developers could use {\tt math.js}, an extensive math library with
support for matrices and big numbers.
\end{itemize}

>>>>

% *** EXTRA HYPHENATION RULES ***
%
\usepackage[english]{babel}
\hyphenation{
  au-tonomous
  non-de-ter-min-is-tic}

@@

Let us see how this system addresses the different challenges
outlined above:
\begin{itemize}
\item {\em Scalability} is provided via the use of a lightweight and
  high-performance, single-threaded, server based in Express.js and
  Node.js. Although this single server is a bottleneck since it
  will eventually saturate, the fact that it runs as a non-blocking single thread
  allows the service of many requests. In fact, a limit in the
  number of simultaneous requests will be reached, but so far it has
  not been found, unlike what we found in our previous systems, DCoR,
  \cite{gecco07:workshop:dcor}, which had a low scalability. 
\item The framework is {\em heterogeneous} since it does not need any
  performance, operating system, or even browser requirement as long
  as JavaScript is enabled: anyone
  visiting the page, even from mobile devices, can load the algorithm.
  In case the browser does not support HTML5 Web workers (e.g., 
  Windows Phone or some version of Android) or their use
  has been disabled, a basic version of {\sf NodIO} can also be used.
\item {\em Fault tolerance} is always an issue, and in this case, the
  single point of failure would be the server: the system, as such,
  would break down if the server fails. However, the individual
  islands in every browser would continue running, and having access
  to just one of them would allow the local algorithm to proceed. In
  fact, the island does not need the server to run: it runs locally if
  needed, with the only exception that it is obviously unable to
  communicate with the rest of the islands.
\item {\em Adaptiveness} is achieved through the autonomous
  operation of every individual island without any synchronization
  mechanism. The islands in the system are, in fact, unaware of each
  other, communicating only through the server. The webpage itself can
  be displayed, using responsive design, in either desktop or mobile
  systems. 
\item Since the algorithm runs in the browser, user's {\em safety} is
  achieved through its sandbox mechanisms. The user is thus assured
  that there is no unsafe access either to their local files or even
  to more resources than the browser should be allotted.
\item Running the algorithm is just a matter of loading the page,
  which makes the operation totally {\em anonymous}. For the same
  reason, {\em ease of use} is optimal, being as easy as simply
  clicking on an URL, available to anyone with access to a browser.
\item {\em Reasonable performance} is not ensured. In fact, we should
  make sure that there is a reasonable amount of clients over which
  the performance achieved is better than what you would obtain in
  your own desktop system. If this is not the case, it is a pointless
  academic exercise.
\end{itemize}

@@

\begin{itemize}
\item {\em Randomize} Both Matlab and Java programs rely on a 
Java Randomizer library for the 
randomization of the shift vectors and matrices used by the function.
The library includes the generation of pseudorandom Gaussians of type
double and uniformly distributed integers. Generators use a single
{\tt long} seed. In order to implement the same functionality in
JavaScript, the library {\tt random-js} was used, as there are inconsistencies 
in the implementation of the standard {\tt Math.random()} between engines
and also their results are non-de\-ter\-min\-is\-tic. 
The library {\tt random-js} implements an algorithm called Mersenne Twister, producing 
consistent results across all JavaScript VMs.
\item {\em Timing functions} In order to compare the performance of 
each implementation, an accurate and consistent way of measuring 
the runtime of functions is needed.
In JavaScript the {\tt Date class} is often used 
to measure execution time, but has the limitation of having a maximum 
resolution given in milliseconds.
In order to measure the runtime intervals with a higher precision two functions
where used: for Node.js we used the native {\tt process.hrtime()} function
which returns a high-resolution measure in a {\tt [seconds, nanoseconds]}
array, and for measuring in the browser the {\tt Per\-for\-mance.now()} function was called,
which uses floating-point numbers with up to microsecond precision. 
A drawback found is that the {\tt Performance.now()} function is only implemented
in Firefox and Chrome brows\-ers. Both functions are independent of the system clock. 
\item {\em Tools} An important consideration when choosing certain 
language to write non-trivial programs is the availability of development tools. While developing
the experiments, we used a comprehensive set of developer tools available for both
browsers and desktop \cite{tilkov2010node}. These tools included
package managers, in this case {\tt npm}, debuggers, logging libraries
such as {\tt winston}, 
network monitors, editors, and even debugging of the multi-threaded execution of Web Workers.
Although the fact that developer tools vary from browser to browser
supposed a drawback, they all have the same functionality. The
JavaScript debugger can be accessed directly from the user menu in
most browsers. 
\item {\em Data Types} JavaScript uses floating point numbers with a 
limited precision of 64 bits that partially implement the functionality of
the {\tt StrictMath} library used by the Java implementation. If more precision
is needed, developers could use {\tt math.js}, an extensive math library with
support for matrices and big numbers.
\end{itemize}
